preciseTAD: A machine learning framework for precise 3D domain boundary prediction at base-level resolution

### 2020-09-09 Manuscript/manuscript.Rmd
Discuss PTBR width, add refs

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. We show that, unlike Arrowhead, *preciseTAD* does not inflate the number of predicted boundaries, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* boundaries were enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143, supporting recent observations that TADs and loops may be generated by similar mechanisms [@hansen2018recent]. Likewise, *preciseTAD* boundaries were more conserved between GM12878 and K562 cell lines, a known feature among the 3D architecture of the human genome, further highlighting their biological significance.

>>>>

Guided by both low-resolution Hi-C data and high-resolution genome annotation data, *preciseTAD* predicts base-level resolution boundaries, alleviating resolution limitations of Hi-C data. However, a natural question is how resolution (width) of boundary regions identified by *preciseTAD* (PTBRs) compares with that of Hi-C data. Our preliminary observations indicate that, under most optimal settings, the width of PTBRs parallels the resolution of Hi-C data. Furthermore, each PTBR is formed by several sub-regions with the probability of being a boundary defined by the threshold $t$ ($t=1$ in the current study). Our results are in line with the emergent view that domain boundaries are dynamic [@chang2020tads] and arise as a consequence of the population average in bulk Hi-C data [@fudenberg2016formation; @Flyamer:2017aa].

We show that, unlike Arrowhead, *preciseTAD* does not inflate the number of predicted boundaries, providing only the most biologically meaningful boundaries that demarcate regions of high inter-chromosomal interactions. *preciseTAD* boundaries predicted using models trained on either TAD or loop boundaries (Arrowhead and Peakachu data) were enriched for known architectural transcription factors including CTCF, RAD21, SMC3, and ZNF143, supporting recent observations that TADs and loops may be generated by similar mechanisms [@hansen2018recent]. Likewise, *preciseTAD* boundaries were more conserved between GM12878 and K562 cell lines, a known feature among the 3D architecture of the human genome, further highlighting their biological significance.
